---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sabre

[![Travis build status](https://travis-ci.org/Nowosad/sabre.svg?branch=master)](https://travis-ci.org/Nowosad/sabre)
[![Coverage status](https://codecov.io/gh/Nowosad/sabre/branch/master/graph/badge.svg)](https://codecov.io/github/Nowosad/sabre?branch=master)

The **sabre** (**S**patial **A**ssociation **B**etween **RE**gionalizations) is an R package for calculating a degree of spatial association between regionalizations or categorical maps.
This package offers support for `sf` spatial objects, and the following methods:

- the V-measure method 
- the MapCurve method (Hargrove et al., 2006)

## Installation

<!-- You can install the released version of `sabre` from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("sabre") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Nowosad/sabre")
```
## Example

We use two simple regionalization, `regions1` and `regions2` to show the basic concept of calculating a degree of spatial association. 

```{r, message=FALSE}
library(sabre)
library(sf)
data("regions1")
data("regions2")
```

The first map, `regions1` consists of four regions of the same shape and size, while the second one, `regions2` has three irregular regions.

```{r, echo=FALSE, fig.height=5, out.width="50%"}
plot(regions1, main = "regions1")
plot(regions2, main = "regions2")
```

The `sabre_calc()` function allows for calculation of a degree of spatial association between regionalizations or categorical maps using the information-theoretical V-measure.
It requires, at least, four arguments:

* `x` - an `sf` object containing the first regionalization
* `x_name` - a name of the column with regions names of the first regionalization
* `y` - an `sf` object containing the second regionalization
* `y_name` -  a name of the column with regions names of the second regionalization

```{r}
regions_vm = sabre_calc(regions1, z, regions2, z)
```

The result is a list with three metrics of spatial association - `V-measure`, `Homogeneity`, `Completeness` - and two `sf` objects with preprocessed input maps - `$map1` and `$map2`.

```{r}
regions_vm
```

Both spatial outputs have two columns. 
The first one contains regions' names/values and the second one (`rih`) describes regions' inhomogeneities.

```{r, eval=FALSE}
plot(regions_vm$map1["rih"], main = "Map1: rih")
plot(regions_vm$map2["rih"], main = "Map2: rih")
```

```{r, warning=FALSE, fig.height=2, fig.height=5, out.width="50%", echo=FALSE}
plot(regions_vm$map1["rih"], main = "Map1: rih"); plot(regions_vm$map2["rih"], main = "Map2: rih")
```

See the package vignette for more examples.

## References

- Rosenberg, Andrew, and Julia Hirschberg. "V-measure: A conditional entropy-based external cluster evaluation measure." Proceedings of the 2007 joint conference on empirical methods in natural language processing and computational natural language learning (EMNLP-CoNLL). 2007.
- Hargrove, William W., Forrest M. Hoffman, and Paul F. Hessburg. "Mapcurves: a quantitative method for comparing categorical maps." Journal of Geographical Systems 8.2 (2006): 187.
