---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sabre

[![Travis build status](https://travis-ci.org/Nowosad/sabre.svg?branch=master)](https://travis-ci.org/Nowosad/sabre)
[![Coverage status](https://codecov.io/gh/Nowosad/sabre/branch/master/graph/badge.svg)](https://codecov.io/github/Nowosad/sabre?branch=master)

The **sabre** (**S**patial **A**ssociation **B**etween **RE**gionalizations) is an R package for calculating a degree of spatial association between regionalizations or categorical maps.
This package offers support for `sf` spatial objects, and the following methods:

- the V-measure method 
- the MapCurve method (Hargrove et al., 2006)

## Installation

<!-- You can install the released version of sabre from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("sabre") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Nowosad/sabre")
```
## Examples

```{r, message=FALSE}
library(sabre)
library(sf)
data("regions1")
data("regions2")
```

```{r, warning=FALSE, echo=FALSE}
library(tmap)
rm1 = tm_shape(regions1) + tm_polygons("z", palette = "Set2") + tm_layout(title = "First regionalization")
rm2 = tm_shape(regions2) + tm_polygons("z", palette = "Set3") + tm_layout(title = "Second regionalization")
tmap_arrange(rm1, rm2) 
```

```{r}
regions_vm = sabre_calc(regions1, z, regions2, z)
regions_vm
```

```{r}
data("eco_us")
```

```{r, echo=FALSE}
em1 = tm_shape(eco_us) + tm_fill("PROVINCE", legend.show = FALSE) + tm_layout("Ecoregions of the United States: Provinces")
em2 = tm_shape(eco_us) + tm_fill("SECTION", legend.show = FALSE) + tm_layout("Ecoregions of the United States: Sections")
tmap_arrange(em1, em2)
```

```{r}
eco_us_vm = sabre_calc(eco_us, PROVINCE, eco_us, SECTION)
eco_us_vm
```

## References

- Rosenberg, Andrew, and Julia Hirschberg. "V-measure: A conditional entropy-based external cluster evaluation measure." Proceedings of the 2007 joint conference on empirical methods in natural language processing and computational natural language learning (EMNLP-CoNLL). 2007.
- Hargrove, William W., Forrest M. Hoffman, and Paul F. Hessburg. "Mapcurves: a quantitative method for comparing categorical maps." Journal of Geographical Systems 8.2 (2006): 187.
